---
title: "Mia_Dust_FunGuilds"
author: "Nat Pombubpa"
date: "Updated on October 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###STEP1: Load all necessary packages for analysis
More information about Phyloseq can be found at the following link: [Phyloseq](https://joey711.github.io/phyloseq/)
If you get error in this step, you probably need to install any packages which causes error.

```{r message=FALSE, warning=FALSE}
library(ape)
library(vegan)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(plyr)
```

###STEP2: Import Mapping file (metadate file)
1.Check mapping file before import to R, R will automatically change sample name that starts with number or contain “-” in sample name. If you get error in this step, you should check sample name first.

2.First column of first row should not start with #, R will not read the first row that starts with #

3.Please make sure that your metadata and FunGuilds SampleID are the same. 

```{r}
meta = read.table("SierraMapIsoF.csv",header=TRUE,row.names=1,sep=",",stringsAsFactors=FALSE)
```

###STEP3: Check if your metadata file has been import successfully and correctly

The output will show a table of your metadata file (mapping file).

*If you do not have header, you might start your first row with #

```{r}
head(meta)
```

```{r}
dim(meta)
```

###STEP4: Construct sample_data-class using imported metadata

```{r}
sampleData <- sample_data(meta)
```

###STEP5: Import FunGuilds table

FunGuilds table from Dust data is “DustFungi.guilds.txt”. This is otu table which also includes Taxon,Trophic Mode, Guild, Growth Morphology etc.)
'FG' contains the entire table, but 'FGotus' is used to select only the otu table part from FunGuilds table (abundance + OTU ID)

```{r}
FG <- read.table("DustFungi.guilds.txt",header=T,sep="\t",row.names=1)
FGotus <- select(FG, -(Taxonomy:Citation.Source))
FGotumat <- as(as.matrix(FGotus), "matrix")
FGOTU = otu_table(FGotumat, taxa_are_rows = TRUE)
```

Entire Data from FunGuilds table (first 6 lines)

```{r}
head(FG)
```

```{r}
dim(FG)
```


Selcted Data from FunGuilds table (first 6 lines)

```{r}
head(FGOTU)
```

```{r}
dim(FGOTU)
```


###STEP6: Import Functional Guilds (Trophic Mode, Guild, Growth.Morphology) as Taxonomy feature

```{r}
FGtaxmat <- select(FG, Confidence.Ranking, Trophic.Mode, Guild, Growth.Morphology, Taxon)
FGtaxmat <- as(as.matrix(FGtaxmat),"matrix")
FGTAX = tax_table(FGtaxmat)
```

```{r}
head(FGTAX)
```

###STEP7: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required.

```{r}
fgps = phyloseq(FGOTU,FGTAX,sampleData)
```

Check phyloseq object

```{r}
fgps
```

###STEP8: Filtering FunGuilds Data
Remove any OTUs which has less than 10 reads
Remove Guilds that have Confidence.Ranking = Possible 
Remove na data
Remove NULL data

Note: you might want to check the entire data before you use these filters.

```{r}
fgps.prune = prune_taxa(taxa_sums(fgps) > 10, fgps)
#fgps.prune.confidence1 = subset_taxa(fgps.prune, Confidence.Ranking!="Possible")
#fgps.prune.no.na = subset_taxa(fgps.prune.confidence1, Trophic.Mode!="-")
#fgps.prune.no.null = subset_taxa(fgps.prune.no.na, Guild!="NULL")
#fgps.prune.no.null = subset_taxa(fgps.prune.no.null, Growth.Morphology!="NULL")
```

Check phyloseq object after filtering

```{r}
fgps.prune.no.null
```

###STEP9: Plot all FunGuilds data from Trophic.Mode, Guild, Growth.Morphology

x = any environmental variable (in this case, I tested with Elevayion)
fill = Functional Guilds that you want to plot (Trophic.Mode, Guild, Growth.Morphology)

```{r fig.height=6, fig.width=10, fig.align="center"}
ggplot(data = psmelt(fgps.prune), mapping = aes_string(x = "Site" ,y = "Abundance", 
                                                               fill = "Trophic.Mode" )) + 
  geom_bar(stat="identity", position="fill") + ggtitle("Dust Fungi Trophic.Mode by Elevation")
```

```{r fig.height=6, fig.width=10, fig.align="center"}
ggplot(data = psmelt(fgps.prune), mapping = aes_string(x = "Elevation" ,y = "Abundance", 
                                                               fill = "Growth.Morphology" )) + 
  geom_bar(stat="identity", position="fill") + ggtitle("Dust Fungi Growth.Morphology by Elevation")
```

```{r fig.height=6, fig.width=12, fig.align="center"}
ggplot(data = psmelt(fgps.prune.no.null), mapping = aes_string(x = "Elevation" ,y = "Abundance", 
                                                               fill = "Guild" )) + 
  geom_bar(stat="identity", position="fill") + ggtitle("Dust Fungi Guild by Elevation")
```

###STEP10: Get the top 100 OTUs and plot only the top abundance

```{r}
ps.prune = prune_taxa(names(sort(taxa_sums(fgps.prune.no.null), TRUE))[1:50], fgps.prune.no.null)
```

```{r fig.height=6, fig.width=12, fig.align="center"}
ggplot(data = psmelt(ps.prune), mapping = aes_string(x = "Elevation" ,y = "Abundance", fill = "Guild" )) + 
  geom_bar(stat="identity", position="fill") + ggtitle("Dust Fungi Guild by Elevation")
```
